### 2024-11-19 MS&E 226_Project Part 1_Code.Rmd
comments added

---
title: "MS&E 226: Project Part 1 Code"
author: "Antra Nakhasi & Sydney Loh"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, warning = FALSE)
knitr::opts_chunk$set(tidy = TRUE, tidy.opts = list(width.cutoff = 60))

options(width = 60) # Set width for knitting code chunks

rm(list = ls()) # Clear environment

# Load necessary libraries
library(tidyverse)
library(dplyr)
library(caret)
library(kknn)
library(reshape2)
library(class)
library(glmnet)
library(e1071)
library(rpart)
library(tibble)
library(randomForest)   
library(xgboost)  
library(quantreg)
library(kernlab)

set.seed(123) # Setting seed for reproducibility
```

# 2: Data Pre-processing

```{r Pre Process, results='hide'}

# load data
df <- read.csv("electricity.csv")

## pre process data

# Convert Revenues.Retail to Revenue.Retail to match singular use of "Revenue" in all other column names
colnames(df)[colnames(df) == "Revenues.Retail"] <- "Revenue.Retail"

# Create binary outcome variable from Uses.Consumed (0 if = 0, 1 otherwise)
df$Uses.Consumed.Binary <- ifelse(df$Uses.Consumed == 0, 0, 1)

# Remove Retail.Total.Sales and Retail.Total.Customers columns (b/c not linearly independent, keep more granular columns)
df <- df[, !names(df) %in% c("Retail.Total.Customers", "Retail.Total.Sales", "Sources.Total", "Uses.Total", 
                             "Retail.Total.Revenue", "Retail.Total.Sales", "Retail.Total.Customers")]

# Convert revenue breakdown columns, Revenue.______, into a categorical variable listing which stream had the greatest share of the total
revenue_cols <- c("Retail.Residential.Revenue", "Retail.Commercial.Revenue", 
                  "Retail.Industrial.Revenue", "Retail.Transportation.Revenue")

# Find the sector with the maximum revenue and remove the revenue columns
df <- df %>%
  rowwise() %>%
  mutate(
    Max.Revenue.Sector = c("Residential", "Commercial", "Industrial", "Transportation")[which.max(c_across(all_of(revenue_cols)))]
  ) %>%
  ungroup() %>%
  select(-all_of(revenue_cols))

northeast_states <- c("CT", "ME", "MA", "NH", "RI", "VT", "NJ", "NY", "PA", "DC")
midwest_states <- c("IL", "IN", "MI", "OH", "WI", "IA", "KS", "MN", "MO", "NE", "ND", "SD")
south_states <- c("DE", "FL", "GA", "MD", "NC", "SC", "VA", "WV", "AL", "KY", "MS", "TN", "AR", "LA", "OK", "TX")
west_states <- c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY", "AK", "CA", "HI", "OR", "WA")

# Assign each state to its corresponding region
df$Utility.Region <- case_when(
  df$Utility.State %in% northeast_states ~ "Northeast",
  df$Utility.State %in% midwest_states ~ "Midwest",
  df$Utility.State %in% south_states ~ "South",
  df$Utility.State %in% west_states ~ "West",
  df$Utility.State == "CN" ~ "Canada",
  TRUE ~ NA_character_ 
)

# Found one NA value where Utility.State was a blank string "" 
# Omit since it was one case at random 
df <- df[-which(is.na(df$Utility.Region)), ]

```

# 3: Hold Out Set

```{r Create Holdout Set, results='hide'}
# Total number of rows in the dataset
n <- nrow(df)

# Sample 20% of the data for the holdout set
holdout_indices <- sample(1:n, size = floor(0.2 * n))

# Create holdout and training sets
holdout <- df[holdout_indices, ]
df <- df[-holdout_indices, ]

# Display the sizes of the datasets
cat("Holdout Set Size:", nrow(holdout), "\n")
cat("Training Set Size:", nrow(df), "\n")
```

# 5: Data Exploration

## Correlations

```{r Correlations with Revenue.Total, fig.show='hide', results='hide'}

# Correlation the continuous outcome: Hours Worked Per Week

# Select relevant columns for analysis
columns_of_interest <- names(df)[!names(df) %in% c("Utility.Number", "Utility.Name", "Utility.State")] # use a different visualization since 1-hot encoding would lead to 50 different states

# Filter data to include only columns of interest
data_subset <- df %>% select(all_of(columns_of_interest))

# Separate numeric and categorical columns
numeric_columns <- names(data_subset)[sapply(data_subset, is.numeric)]
categorical_columns <- setdiff(columns_of_interest, numeric_columns)

# One-hot encode categorical variables
dummies <- dummyVars(" ~ .", data = data_subset[categorical_columns])
encoded_categorical_data <- predict(dummies, newdata = data_subset) %>% as.data.frame()

# Combine numeric data with one-hot encoded categorical data
encoded_data <- cbind(data_subset[numeric_columns], encoded_categorical_data)

# Calculate all correlations
correlations <- cor(encoded_data, use = "complete.obs")

# Calculate correlations with "Hours_Worked_Per_Week"
correlations_with_revenue <- correlations["Revenue.Total",]

# Sort and display correlations
sorted_correlations_revenue <- sort(correlations_with_revenue, decreasing = TRUE)
# Create a data frame of sorted correlations for better readability
sorted_correlations_revenue <- data.frame(
  Variable = names(sorted_correlations_revenue),
  Correlation_with_revenue = sorted_correlations_revenue
)

# Display the data frame in a readable format
print(sorted_correlations_revenue, row.names = FALSE)

top_correlated_vars <- head(sorted_correlations_revenue$Variable[-1], 5)
# Generate scatterplots for each of the top correlated variables with Revenue.Total
for (var in top_correlated_vars) {
  p <- ggplot(df, aes_string(x = var, y = "Revenue.Total")) +
    geom_point(alpha = 0.6, color = "blue") +
    labs(title = paste("Scatterplot of", var, "vs Revenue.Total"),
         x = var, y = "Revenue.Total") +
    theme_minimal() +
    geom_smooth(method = "lm", color = "red", se = FALSE)

  print(p)  # Explicitly print each plot
}

ggplot(sorted_correlations_revenue, aes(x = reorder(Variable, Correlation_with_revenue), 
                                             y = Correlation_with_revenue)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip coordinates for better readability
  labs(title = "Correlation with Revenue.Total", 
       x = "Variables", 
       y = "Correlation Coefficient") +
  theme_minimal()
```

```{r Correlations with Uses.Consumed.Binary, fig.show='hide', results='hide'}
# Calculate correlations with binary outcome variable
correlations_with_uses_consumed <- correlations["Uses.Consumed.Binary",]

# Sort and display correlations
sorted_correlations_uses_consumed <- sort(correlations_with_uses_consumed, decreasing = TRUE)

# Create a data frame of sorted correlations for better readability
sorted_correlations_uses_consumed <- data.frame(
  Variable = names(sorted_correlations_uses_consumed),
  Correlation_with_Uses_Consumed = sorted_correlations_uses_consumed
)

# Display the data frame in a readable format
print(sorted_correlations_uses_consumed, row.names = FALSE)

ggplot(sorted_correlations_uses_consumed, aes(x = reorder(Variable, Correlation_with_Uses_Consumed), 
                                             y = Correlation_with_Uses_Consumed)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip coordinates for better readability
  labs(title = "Correlation with Revenue.Total", 
       x = "Variables", 
       y = "Correlation Coefficient") +
  theme_minimal()

```

```{r correlation of covariates with each other, fig.show='hide', results='hide'}

# Create a list of numeric columns, not including the binary outcome variable 
cont_cov_columns <- numeric_columns[numeric_columns != "Uses.Consumed.Binary"]

# Calculate correlations for continuous variables
correlations_cont <- cor(data_subset[cont_cov_columns], use = "complete.obs")

# Melt the correlation matrix for ggplot2
cor<- melt(correlations_cont)

# Create the heatmap
ggplot(data = cor, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name="Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 10, hjust = 1)) +
  coord_fixed()

```

## Subgroups in data

```{r means across categories, fig.show='hide', results='hide'}

# Calculate means for utility.state
state_means <- df %>%
  summarise(across(-c(Utility.Number, Utility.Name, Utility.Type, Max.Revenue.Sector, Uses.Consumed.Binary, Utility.Region),
                   ~ round(mean(.x, na.rm = TRUE), 5)),
            .by = Utility.State)

# Calculate means for utility.type
type_means <- df %>%
  summarise(across(-c(Utility.Number, Utility.Name, Utility.State, Max.Revenue.Sector, Uses.Consumed.Binary, Utility.Region),
                   ~ round(mean(.x, na.rm = TRUE), 5)),
            .by = Utility.Type)

region_means <- df %>%
  summarise(across(-c(Utility.Number, Utility.Name, Utility.State, Max.Revenue.Sector, Uses.Consumed.Binary, Utility.Type),
                   ~ round(mean(.x, na.rm = TRUE), 5)),
            .by = Utility.Region)

View(state_means)
View(type_means)
View(region_means)
variables_to_plot <- names(state_means)[-1]  # Exclude Utility.State
plot_list <- list()
```

```{r Utility.State plots, fig.show='hide', results='hide'}
# Loop through each variable and create a plot
for (var in variables_to_plot) {
  p <- ggplot(state_means, aes_string(x = "Utility.State", y = var)) +
    geom_bar(stat = "identity", fill = "green") +
    coord_flip() +
    labs(title = paste("Mean", var, "by Utility State"),
         x = "Utility State",
         y = paste("Mean", var)) +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 6, angle = 0),
          axis.title.y = element_text(size = 10),
          axis.title.x = element_text(size = 10),
          plot.title = element_text(size = 12, face = "bold"))

  plot_list[[var]] <- p
}

for (plot in plot_list) {
  print(plot)
}

```

```{r Utility.Type plots, fig.show='hide', results='hide'}
# Create plots for utility type means
variables_to_plot_type <- names(type_means)[-1]  # Exclude Utility.Type
plot_list_type <- list()

# Loop through each variable and create a plot for utility type
for (var in variables_to_plot_type) {
  p <- ggplot(type_means, aes_string(x = "Utility.Type", y = var)) +
    geom_bar(stat = "identity", fill = "coral") +
    coord_flip() +
    labs(title = paste("Mean", var, "by Utility Type"),
         x = "Utility Type",
         y = paste("Mean", var)) +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 6, angle = 0),
          axis.title.y = element_text(size = 10),
          axis.title.x = element_text(size = 10),
          plot.title = element_text(size = 12, face = "bold"))

  plot_list_type[[var]] <- p
}

for (plot in plot_list_type) {
  print(plot)
}

```

```{r Utility.Region Plots, fig.show='hide', results='hide'}
# Define the list of variables to plot, excluding the grouping column (Utility.Region)
variables_to_plot <- names(region_means)[-1]

# Initialize an empty list to store the plots
plot_list <- list()

# Loop through each variable in region_means and create a plot
for (var in variables_to_plot) {
  p <- ggplot(region_means, aes_string(x = "Utility.Region", y = var)) +
    geom_bar(stat = "identity", fill = "green") +
    coord_flip() +
    labs(title = paste("Mean", var, "by Utility Region"),
         x = "Utility Region",
         y = paste("Mean", var)) +
    theme_minimal() +
    theme(axis.text.y = element_text(size = 6, angle = 0),
          axis.title.y = element_text(size = 10),
          axis.title.x = element_text(size = 10),
          plot.title = element_text(size = 12, face = "bold"))

  # Add each plot to the list
  plot_list[[var]] <- p
}

# Print each plot for region_means
for (plot in plot_list) {
  print(plot)
}
```

# 6: Regression

## Data Transformations

```{r data transformations}
# set seed
set.seed(123)

### Choosing prediction error metric
# Calculate summary statistics for revenue data
mean_revenue <- mean(df$Revenue.Total)
median_revenue <- median(df$Revenue.Total)
std_dev_revenue <- sd(df$Revenue.Total)
iqr_revenue <- IQR(df$Revenue.Total)

# Print summary statistics
print(paste("Mean Revenue:", mean_revenue))
print(paste("Median Revenue:", median_revenue))
print(paste("Standard Deviation:", std_dev_revenue))
print(paste("IQR:", iqr_revenue))

# Define the columns to include in df_pred
column_names <- c(
  "Demand.Summer.Peak", "Demand.Winter.Peak", "Sources.Generation", "Sources.Purchased",
  "Sources.Other", "Uses.Retail", "Uses.Resale", "Uses.No.Charge", "Uses.Consumed", "Uses.Losses",
  "Retail.Residential.Customers", "Retail.Commercial.Customers", "Retail.Industrial.Customers",
  "Retail.Transportation.Customers", "Utility.Region", "Max.Revenue.Sector", "Revenue.Total", "Utility.Type"
)

# Create df_pred with only the specified columns from df
df_pred <- df[, column_names]

# # Display df_pred
# df_pred

### Applying data transformations
# Applying log transformation to the Revenue.Total column
df_pred$Log_Revenue_Total <- log1p(df_pred$Revenue.Total)

# Histogram
ggplot(df_pred, aes(x = Revenue.Total)) + 
  geom_histogram(binwidth = 5000, fill = "skyblue", color = "black", alpha = 0.7) + 
  labs(title = "Histogram of Revenue Total", x = "Revenue Total", y = "Frequency") +
  theme_minimal()

# Histogram of log-transformed Revenue.Total (after scaling)
ggplot(df_pred, aes(x = Log_Revenue_Total)) + 
  geom_histogram(binwidth = 0.1, fill = "lightgreen", color = "black", alpha = 0.7) + 
  labs(title = "Histogram of Log-transformed Revenue Total", 
       x = "Log(Revenue Total)", 
       y = "Frequency") +
  theme_minimal()

```

## Baseline Model

```{r Baseline Model}
### Baseline Model
# Fit OLS Model (Linear Regression) without cross-validation
ols_model <- lm(Log_Revenue_Total ~ ., 
                data = df_pred)  # Fit the model on the entire dataset

# Print the summary of the OLS model
summary(ols_model)

# Get the residuals to calculate RMSE
predictions <- predict(ols_model, newdata = df_pred)
residuals <- df_pred$Log_Revenue_Total - predictions

# Calculate RMSE (Root Mean Squared Error)
rmse <- sqrt(mean(residuals^2))
cat(paste("OLS Model RMSE:", rmse, "\n"))
```

## Cross Validation setup

```{r cross validation}
### Cross Validation
# Set up cross-validation control with 10 folds
train_control <- trainControl(method = "cv", 
                              number = 10, 
                              summaryFunction = defaultSummary,
                              search = "grid",
                              savePredictions = "all")
```

## Model 1: Ridge Regression

```{r Model 1: Ridge Regression}
### Ridge Regression
ridge_model <- train(Log_Revenue_Total ~ ., 
                     data = df_pred, 
                     method = "glmnet",  # glmnet is used for regularized regression
                     trControl = train_control,
                     metric = "RMSE", 
                     preProcess = "scale",  # Standardize the data
                     tuneGrid = expand.grid(alpha = 0, lambda = seq(0, 1, length = 100))) # alpha=0 for Ridge

# Print CV error 
cat("Ridge Regression CV RMSE for each fold:\n")
print(ridge_model$resample$RMSE)

ridge_model_rmse <- min(ridge_model$results$RMSE)
cat(paste("Ridge Regression RMSE:", ridge_model_rmse, "\n"))

final_lambda_ridge <- ridge_model$bestTune$lambda
cat(paste("Final lambda value for Ridge regression:", final_lambda_ridge, "\n"))

```

## Model 2: Lasso Regression

```{r Model 2: Lasso Regression}
### Lasso Regression
# Fit Lasso regression using glmnet with tuning of lambda and data standardization
lasso_model <- train(Log_Revenue_Total ~ ., 
                     data = df_pred, 
                     method = "glmnet",  # glmnet for regularized regression
                     trControl = train_control,
                     metric = "RMSE",  # RMSE for model evaluation
                     preProcess = "scale",  # Standardize the data
                     tuneGrid = expand.grid(alpha = 1, lambda = seq(0, 1, length = 100)))  # Lasso: alpha = 1

# Print CV error (RMSE for each fold)
cat("Lasso Regression CV RMSE for each fold:\n")
print(lasso_model$resample$RMSE)

# Extract and print the best RMSE for Lasso Regression
lasso_model_rmse <- min(lasso_model$results$RMSE)
cat(paste("Lasso Regression Best RMSE:", lasso_model_rmse, "\n"))

# Extract and print the best lambda value for Lasso regression
final_lambda_lasso <- lasso_model$bestTune$lambda
cat(paste("Final lambda value for Lasso regression:", final_lambda_lasso, "\n"))

```

## Model 3: Random Forest

```{r Model 3: Random Forest}
### Random Forest
# Fit a Random Forest model
rf_model <- train(Log_Revenue_Total ~ ., 
                  data = df_pred, 
                  method = "rf",  # Random Forest regression
                  trControl = train_control,
                  metric = "RMSE",
                  tuneGrid = expand.grid(mtry = sqrt(ncol(df_pred)))) # Hyperparameter tuning for mtry

# Print CV error
cat("Random Forest CV RMSE for each fold:\n")
print(rf_model$resample$RMSE)

rf_model_rmse <- min(rf_model$results$RMSE)
cat(paste("Random Forest RMSE:", rf_model_rmse, "\n"))
```

# 7: Classification

## Data Cleaning

```{r clean data for building classifiers}

# Select specific columns from the dataframe based on positive correlation > 0.1 or negative correlation <-0.1
df_clean <- df[, c("Uses.Consumed.Binary", # Binary outcome variable
                   "Utility.Type", # Positively or negatively correlated with outcome, depending on factor level 
                   "Demand.Winter.Peak", # Positively correlated with outcome
                   "Demand.Summer.Peak", # Positively correlated with outcome
                   "Max.Revenue.Sector", # Negatively correlated with outcome
                   "Revenue.Retail", # Positively correlated with outcome
                   grep("^Retail\\.", names(df), value = TRUE))] # Positively correlated with outcome

# Loop through each column in the dataset and apply the factor transformation to each non-numeric (categorical) variable
df_clean[] <- lapply(df_clean, function(x) {
  if (!is.numeric(x)) {
    factor(x, levels = unique(x))
  } else {
    x
  }
})

df_clean$Uses.Consumed.Binary <- factor(df_clean$Uses.Consumed.Binary, levels = c("0", "1"))
levels(df_clean$Uses.Consumed.Binary) <- make.names(levels(df_clean$Uses.Consumed.Binary))

# Create a copy of the original dataframe to store the scaled data
df_clean_scaled <- df_clean

# Identify numeric columns, excluding 'Uses.Consumed.Binary'
numeric_cols <- sapply(df_clean, is.numeric) & names(df_clean) != "Uses.Consumed.Binary"

# Apply scale() to the selected numeric columns
df_clean_scaled[numeric_cols] <- scale(df_clean[numeric_cols])

```

## Classifier 1: K-NN

```{r Classifier 1: K-NN}

# Split the data into features and target variable
X <- df_clean_scaled[, !names(df_clean) %in% "Uses.Consumed.Binary"] # continuous outcome variable and linear combination of other columns
# Step 1: Prepare the data
# Identify numeric and categorical columns
numeric_cols <- sapply(X, is.numeric)
categorical_cols <- sapply(X, function(x) is.factor(x) | is.character(x))

# Create dummy variables for categorical features
dummy <- dummyVars(" ~ .", data = X[, categorical_cols])
categorical_dummy <- predict(dummy, newdata = X[, categorical_cols])

# Combine numeric and dummy-coded categorical variables
X <- cbind(X[, numeric_cols], categorical_dummy)

# Ensure the target variable is a factor
y <- df_clean_scaled$Uses.Consumed.Binary

# Step 2: Define a function to find the best k using cross-validation

# Define a custom summary function for 0-1 loss
zero_one_loss <- function(data, lev = NULL, model = NULL) {
  loss <- mean(data$obs != data$pred)
  c(loss = loss)
}

# Find_best_k function
find_best_k <- function(X, y, k_range, distance_metric) {
  set.seed(123)

  # Define custom summary function for 0-1 loss
  zero_one_loss <- function(data, lev = NULL, model = NULL) {
    loss <- mean(data$obs != data$pred)
    c(loss = loss)
  }

  control <- trainControl(method = "cv",
                          number = 10,
                          summaryFunction = zero_one_loss)

  grid <- expand.grid(k = k_range)

  knn_model <- train(x = X, y = y,
                     method = "knn",
                     tuneGrid = grid,
                     trControl = control,
                     metric = "loss",
                     maximize = FALSE)  # We want to minimize the loss

  best_k <- knn_model$bestTune$k
  best_loss <- min(knn_model$results$loss)

  return(list(best_k = best_k, mean_0_1_loss = best_loss))
}

# Step 3: Find the best k
k_range <- seq(1, 20, by = 2)
best_k <- find_best_k(X, y, k_range)

# Step 4: Train the final k-NN model with the best k
final_model <- knn(train = X, test = X, cl = y, k = best_k[[1]])

# Save mean 0-1 loss for best k
Err_CV_knn <- best_k[[2]]

# Print results
print(paste("Best k:", best_k[[1]]))
print(paste("0-1 Loss:", round(Err_CV_knn, 4)))

```

## Classifier 2: Logistic Regression with Feature Selection

```{r Classfier 2: Logistic Regression with Feature Selection}

# Create model formula
formula <- as.formula("Uses.Consumed.Binary ~ Utility.Type + Max.Revenue.Sector + Demand.Winter.Peak + Demand.Summer.Peak + Revenue.Retail + Revenue.Retail:Max.Revenue.Sector")

# Set up cross-validation for classification
ctrl <- trainControl(
  method = "cv",
  number = 10,
  classProbs = TRUE,
  summaryFunction = defaultSummary
)

# Define grid of lambda values to tune for regularization
tune_grid <- expand.grid(
  alpha = 0,       # Set alpha = 0 for ridge regression (L2 regularization)
  lambda = seq(0.001, 0.1, by = 0.01) # Define a sequence of lambda values
)

# Train the logistic regression model with cross-validation and lambda tuning
model_cv <- train(
  formula,
  data = df_clean_scaled,
  method = "glmnet",
  family = "binomial",
  trControl = ctrl,
  metric = "Accuracy",
  tuneGrid = tune_grid
)

# Calculate 0-1 loss for each fold
zero_one_loss <- 1 - model_cv$resample$Accuracy

# Calculate mean 0-1 loss across folds
Err_CV_logr <- mean(zero_one_loss)

# Train final model on all data with the best lambda
final_model <- glmnet(
  as.matrix(df_clean_scaled[, all.vars(formula)[-1]]), 
  df_clean_scaled$Uses.Consumed.Binary, 
  alpha = 0,
  lambda = model_cv$bestTune$lambda,
  family = "binomial"
)

# Print results
print(paste("0-1 Loss:", round(Err_CV_logr, 4)))
print(paste("Best lambda:", model_cv$bestTune$lambda))

```

## Classifier 3: Decision Tree

```{r Classifier 3: Decision Tree}

set.seed(123)  # For reproducibility

# Create model formula
formula <- as.formula("Uses.Consumed.Binary ~ Utility.Type + Max.Revenue.Sector + Demand.Winter.Peak + Demand.Summer.Peak + Revenue.Retail + Revenue.Retail:Max.Revenue.Sector")

# Define the hyperparameter grid
hyper_grid <- expand.grid(cp = seq(0.01, 0.1, by = 0.01))

# Set up 10-fold cross-validation with Accuracy as the selection metric
ctrl <- trainControl(
  method = "cv",
  number = 10
)

# Perform grid search with cross-validation
model_grid <- train(
  formula,
  data = df_clean_scaled,
  method = "rpart",
  trControl = ctrl,
  tuneGrid = hyper_grid,
  metric = "Accuracy"
)

# Get the best model
best_model <- model_grid$finalModel

# Calculate the average cross-validation error (1 - Accuracy)
Err_CV_dt <- 1 - max(model_grid$results$Accuracy)

# Display the average cross-validation error
print(paste("0-1 Loss:", round(Err_CV_dt, 4)))
# Display the best model parameters
print(paste("cp of chosen model:", model_grid$bestTune))

```

## Comparing Classifiers

```{r compare classifiers, include=TRUE, echo=FALSE, results='asis'}

Err_baseline <- round(sum(df$Uses.Consumed.Binary)/nrow(df), 4)

Errs <- c(Baseline = Err_baseline, Knn = round(Err_CV_knn, 4), Logr = round(Err_CV_logr, 4), DT = round(Err_CV_dt, 4))

# Convert to a tibble
Errs_table <- tibble(Model = c("Baseline", "K-Nearest Neighbors", "Logistic Regression (Ridge)", "Decision Tree"), Error = sprintf("%.4f", Errs))

# Print the table
knitr::kable(Errs_table, col.names = c("Model", "Prediction Error (Mean 0-1 Loss)"))

print(Errs_table)
```